Practical Effects of Integrating Temperature with Strang Split Reactions

### 2021-03-10 strang/paper.tex
Add Gadget cite
Add Raskin cite

There are several different approximations that people make in the
literature to this operator-split reacting system, in order to make
integrating the reaction system computationally less expensive:
\begin{itemize}
\item Evolve $(X_k)$ only.  This neglects temerature evolution
  completely in the burn, only evolving Eq.~\ref{eq:species}.  This is
  the method used in \citet{flash}. \MarginPar{other codes?}

\item Evolve $(X_k, T)$.  To avoid expensive equation of state calls
  in getting the specific heat, we can optionally ``freeze'' the value
  of $c_v$ at the start of integration to avoid potentially expensive
  equation of state calls.  This was discussed in \cite{Bell:2004} and
  until recently was the default method in \castro.

\item Evolve $(X_k, e)$, and get $T$ from $e$ using the equation of
  state.  This was discussed in \citet{fma} and is the current default
  method in \castro.
\end{itemize}
For all of these cases, density is constant during the reaction operation.
Depending on how vigorous the burning is and how much the temperature
changes during a hydrodynamic timestep, one or more of these
methods may be reasonable.  For explosive reactions, we expect
that evolving the full system will be required.  The goal of this
note is to try to quantify the convergence of a reacting hydrodynamics
code with these different approximations.

>>>>

There are several different approximations that people make in the
literature to this operator-split reacting system, in order to make
integrating the reaction system computationally less expensive:
\begin{itemize}
\item Evolve $(X_k)$ only.  This neglects temperature evolution
  completely in the burn, only evolving Eq.~\ref{eq:species}.  This is
  the method used in \citet{flash}. \MarginPar{other codes?}

\item Evolve $(X_k, T)$.   This is used in \cite{Pakmor:2012}.  To avoid
  expensive equation of state calls
  in getting the specific heat, we can optionally ``freeze'' the value
  of $c_v$ at the start of integration to avoid potentially expensive
  equation of state calls.  This was discussed in \cite{Bell:2004} and
  until recently was the default method in \castro.

\item Evolve $(X_k, e)$, and get $T$ from $e$ using the equation of
  state.  This was discussed in \citet{fma} and is the current default
  method in \castro.
\end{itemize}
\cite{Raskin:2010} also propose a hybrid system where the first approach is
used in most cases, switching to the second approach only near NSE.
For all of these cases, density is constant during the reaction operation.
Depending on how vigorous the burning is and how much the temperature
changes during a hydrodynamic timestep, one or more of these
methods may be reasonable.  For explosive reactions, we expect
that evolving the full system will be required.  The goal of this
note is to try to quantify the convergence of a reacting hydrodynamics
code with these different approximations.
